Investigating soundscapes perception through acoustic scenes simulation

### 2018-07-15 lafaySimulatedScenePerception.tex
Add GL revisions

@@

In order to account for the possibility that the morphological differences between those two classes may have some important consequences on the perception of the scenes, the simulation tool \emph{Simscene} follows this distinction and considers two distinct sound databases: one with classes of events only and the other with classes of textures only. These two types of sound classes also have specific scheduling procedures during the simulation process.

@@

According to the Auditory Scene Analysis (ASA) theory, the brain processes separately events emerging from the different sources of a sound environment~\cite{bregman1994auditory}. Several auditory streams are created, one for each sequence of events emitted by the same source \cite{carlyon2004brain}, while events that cannot be isolated are grouped into a same stream.

@@

Maffiolo~\cite{maffiolo_caracterisation_1999} distinguishes two separated categorization processes, either of which is triggered depending on the listener's ability to identify sound events. She shows that those processes lead to two abstract cognitive categories respectively termed "event sequences" and "amorphous sequences".

@@

Event sequences are composed of salient events that can easily be recognized, such as \emph{car start} or \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On the contrary, amorphous sequences are sound environments where distinct events cannot be readily identified, such as \emph{traffic hubhub}. They result from a holistic analysis based on global acoustic features.

@@

Concerning sound textures, \ie~sounds that have stable characteristics over time,~\cite{mcdermott2011sound,mcdermott2013summary} demonstrates that the human brain can opt for an abstract, statistical representation of the perceived information, discarding precise physical properties of the sound~\cite{nelken2013ear}.

@@

In the light of those results, it appears that the processing of auditory information comprises some sort of decision concerning the nature of the stimuli~\cite{nelken2013ear,mcdermott2013summary}. We thus consider in this study the soundscape as a ``\,skeleton of events on a bed of textures\,'' as coined in~\cite{nelken2013ear}.

@@

\begin{itemize}
\item \emph{experiment 1.a (simulation)}: during this experiment, subjects are asked to simulate urban sound environments using \emph{Simscene} (see Section~\ref{sec:simulator}). Each of them has to create two sound environments: one ideal~/ pleasant, and the other non ideal~/ unpleasant.
\item \emph{experiment 2.a (evaluation)}: after the simulation phase, only a binary information on the pleasantness property of the respective scenes is available: respectively ideal (i) or non ideal (ni). Furthermore, this information is given by the creator of the scene. The second experimental step aims at investigating more deeply and more broadly our knowledge on the pleasantness of the simulated scenes. For that, a second group of subjects is asked to evaluate the pleasantness of each scene produced during (1.a), on a semantic scale. This experiment has two goals:
\begin{enumerate}
\item to evaluate more precisely the respective influence of the various sources composing the scenes on the pleasantness  (ideal or non ideal) thanks to a finer quantification of the pleasantness of the scene;
\item to detect the presence of \emph{outliers} or ambiguous scenes. Indeed, throughout our analyses, the predefined hedonic properties of the scenes (ideal or not) are used as reference. We thus need to ensure beforehand that no ambiguity exists between extreme cases of ideal and non ideal scenes, \ie~that the least pleasant ideal scene remains more highly rated on average than the most pleasant non ideal scene.
\end{enumerate}
\end{itemize}

@@

We have shown that, from a qualitative point of view, the composition of the scenes in terms of sound sources differs between i- or non ideal scenes. We now investigate whether some of the sound classes are specific to a given environment. For that purpose, the V-test detailed in  Section~\ref{sec:xp1_dataAna} is considered separately for each abstraction level. Results are presented in Table~\ref{tab:markers}.

@@

Considering $E$ and $T$ separately, it appears that: 1) the rank 5-precision obtained with $E$ is similar to the one obtained with $ET$ ; 2) the precision obtained with $T$ is always lower than the one obtained with $E$, by $10\%$ to $15\%$. Those results indicate that the semantic information that is discriminative is mostly carried by the events. Those results are in line with results of~\cite{maffiolo_caracterisation_1999}. As discussed in Section~\ref{sec:simscene_sampleDataSet}, it seems that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie~by identifying the sources.

@@

To do so, the correlations between $\mathcal{A}_{scene}$ and the sound levels are evaluated. In this section, the sound levels are computed by taking into account only the previously identified sound markers. We define $X_m$, the $X$ feature computed by taking into account only the sound markers, and $X_b$, the $X$ feature computed by taking into account all the sound classes, except the sound markers. When the feature characterizes an ideal scene (resp. non ideal scene), only the markers identified for the ideal scenes (resp. non ideal scenes) are considered. We henceforth call \emph{ideal markers} and \emph{non ideal markers} the two types of markers. Results are shown on Table~\ref{tab:corrMarkers}.

@@

Notwithstanding the presence or absence of markers, the km-scenes and rm-scenes are exactly the same.

@@

In order to create rm-scenes with still some sound diversity and no  absence of sound activity within a long period of the scene, only the sound classes of events of the first level of abstraction are removed, see Table~\ref{tab:markers}. Those classes are:

@@

For each subject, the scenes are presented in a random order. The first 10 scenes allow the subject to calibrate their scores. Those calibration scenes are  5 ideal scenes with markers and 5 nideal scenes with markers. These first 10 scenes are replayed at the end of the experiment, and only the scores given at the second occurrence are taken into account.

@@

The aim here is to validate the hypothesis that the removal of ideal markers and non ideal markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{subject}$ as a dependent variable, and as independent variables, the type of environment (i/ni) and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to evaluate whether there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure. %\jfp{l'analyse est inutile a mon avis, il n'y a que 2 niveaux pour chaque facteur de l'ANOVA.} \gl{2 niveaux par facteur, mais 4 niveaux en tout}

@@

Let us consider $\mathcal{A}_{subject}$ for the km-scenes (Figure~\ref{fig:xp2_Asubject}). Close inspection of the answers show that one subject's judgments strongly differ from the others'. This subject evaluated positively close to half of the non ideal scenes with markers (see Annex~\ref{app:xp2}, subject 7), and gave a score above 0 for 58\% of the nideal scenes with markers, contrary to an average of 11\% for the other subjects.

@@

In this section, we study the scores given by the subject while listening to the several types of scenes, namely the ideal scenes with markers, nideal scenes with markers, ideal scenes without markers and non ideal scenes without markers (Figure~\ref{fig:xp2_Ascene}). The repeated measure ANOVA applied to $\mathcal{A}_{subject}$ shows a significant effect of the type environment (i/ni: $F[1,10]=175$, $p<0.01$), of the presence/absence of markers (km/rm: $F[1,10]=7$, $p<0.05$), and of the interaction between those two factors ($F[1,10]=67$, $p<0.01$).

@@

The \emph{post hoc} analysis exhibits significant differences between all groups of observations, notably between the i/km- and the i/rm scenes ($p<0.05$) as well as between the ni/km- and the non ideal scenes without markers ($p<0.01$).

@@

The significant interaction shows that the type of environment influences the effect of the presence/absence of the markers. Indeed, the average difference between km-scenes and rm-scenes is larger for the non ideal scenes ($1.1$) than for the ideal scenes ($0.5$).

>>>>

It indeed appears that the processing of auditory information comprises some sort of decision concerning the nature of the stimuli.

\cite{maffiolo_caracterisation_1999} distinguishes two separated categorization processes, either of which is triggered depending on the listener's ability to identify sound events. She shows that those processes lead to two abstract cognitive categories respectively termed "event sequences" and "amorphous sequences". Event sequences are composed of salient events that can easily be recognized, such as \emph{car start} or \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On the contrary, amorphous sequences are sound environments where distinct events cannot be readily identified, such as \emph{traffic hubhub}. They result from a holistic analysis based on global acoustic features.

Concerning sound textures, \ie~sounds that have stable characteristics over time, \cite{mcdermott2011sound,mcdermott2013summary} demonstrates that the human brain can opt for an abstract, statistical representation of the perceived information, discarding precise physical properties of the sound.

In order to account for the possibility that the morphological differences between soud events and sound textures may have some important consequences on the perception of the scenes, the simulation tool \emph{Simscene} follows this distinction and considers two distinct sound databases: one with classes of events only and the other with classes of textures only. These two types of sound classes also have specific scheduling procedures during the simulation process.

We consider in this study the soundscape as a ``\,skeleton of events on a bed of textures\,'' as coined by~\cite{nelken2013ear}.

@@

@@

%According to the Auditory Scene Analysis (ASA) theory, the brain processes separately events emerging from the different sources of a sound environment~\cite{bregman1994auditory}. Several auditory streams are created, one for each sequence of events emitted by the same source \cite{carlyon2004brain}, while events that cannot be isolated are grouped into a same stream.

@@

@@

@@

%Concerning sound textures, \ie~sounds that have stable characteristics over time,~\cite{mcdermott2011sound,mcdermott2013summary} demonstrates that the human brain can opt for an abstract, statistical representation of the perceived information, discarding precise physical properties of the sound~\cite{nelken2013ear}.

@@

% mcdermott2013summary

@@

\begin{itemize}
\item \emph{experiment 1.a (simulation)}: during this experiment, subjects are asked to simulate urban sound environments using \emph{Simscene} (see Section~\ref{sec:simulator}). Each of them has to create two sound environments: one ideal~/ pleasant, and the other non ideal~/ unpleasant.
\item \emph{experiment 2.a (evaluation)}: after the simulation phase, only a binary information on the pleasantness property of the respective scenes is available: respectively ideal (i) or non ideal (ni). Furthermore, this information is given by the creator of the scene. The second experimental step aims at investigating more deeply and more broadly our knowledge on the pleasantness of the simulated scenes. For that, a second group of subjects is asked to evaluate the pleasantness of each scene produced during (1.a), on a semantic scale. This experiment has two goals:
\begin{enumerate}
\item to evaluate more precisely the respective influence of the various sources composing the scenes on the pleasantness  (ideal or non ideal) thanks to a finer quantification of the pleasantness of the scene;
\item to detect the presence of \emph{outliers} or ambiguous scenes. Indeed, throughout our analyses, the predefined hedonic properties of the scenes (ideal or non ideal) are used as reference. We thus need to ensure beforehand that no ambiguity exists between extreme cases of ideal and non ideal scenes, \ie~that the least pleasant ideal scene remains more highly rated on average than the most pleasant non ideal scene.
\end{enumerate}
\end{itemize}

@@

We have shown that, from a qualitative point of view, the composition of the scenes in terms of sound sources differs between ideal or non ideal scenes. We now investigate whether some of the sound classes are specific to a given environment. For that purpose, the V-test detailed in  Section~\ref{sec:xp1_dataAna} is considered separately for each abstraction level. Results are presented in Table~\ref{tab:markers}.

@@

Considering $E$ and $T$ separately, it appears that: 1) the rank 5-precision obtained with $E$ is similar to the one obtained with $ET$ ; 2) the rank 5-precision obtained with $T$ is always lower than the one obtained with $E$, by $10\%$ to $15\%$. Those results indicate that the semantic information that is discriminative is mostly carried by the events. Those results are in line with results of~\cite{maffiolo_caracterisation_1999}. As discussed in Section~\ref{sec:simscene_sampleDataSet}, it seems that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie~by identifying the sources.

@@

To do so, the correlations between $\mathcal{A}_{scene}$ and the sound levels are evaluated. In this section, the sound levels are computed by taking into account only the previously identified sound markers. We define $L_m$ (resp. $L(E)_m$ and $L(T)_m$), the sound level computed by taking into account only the sound markers, and $L_b$ (resp. $L(E)_b$ and $L(T)_b$), the sound level computed by taking into account all the sound classes, except the sound markers. When the feature characterizes an ideal scene (resp. non ideal scene), only the markers identified for the ideal scenes (resp. non ideal scenes) are considered. We henceforth call \emph{ideal markers} and \emph{non ideal markers} the two types of markers. Results are shown on Table~\ref{tab:corrMarkers}.

@@

Notwithstanding the presence or absence of markers, the scenes with and without markers are exactly the same.

@@

In order to create the scenes without markers with still some sound diversity and no absence of sound activity within a long period of the scene, only the sound classes of events of the first level of abstraction are removed, see Table~\ref{tab:markers}. Those classes are:

@@

For each subject, the scenes are presented in a random order. The first 10 scenes allow the subject to calibrate their scores. Those calibration scenes are  5 ideal scenes with markers and 5 non ideal scenes with markers. These first 10 scenes are replayed at the end of the experiment, and only the scores given at the second occurrence are taken into account.

@@

The aim here is to validate the hypothesis that the removal of ideal markers and non ideal markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{subject}$ as a dependent variable, and as independent variables, the type of environment (ideal or non ideal) and the presence/absence of markers. As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to evaluate whether there exist a significant difference of perceived pleasantness. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: ideal or non ideal,  marker: with or without), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure. %\jfp{l'analyse est inutile a mon avis, il n'y a que 2 niveaux pour chaque facteur de l'ANOVA.} \gl{2 niveaux par facteur, mais 4 niveaux en tout}

@@

Let us consider $\mathcal{A}_{subject}$ for the scenes with markers (Figure~\ref{fig:xp2_Asubject}). Close inspection of the answers show that one subject's judgments strongly differ from the others'. This subject evaluated positively close to half of the non ideal scenes with markers (see Annex~\ref{app:xp2}, subject 7), and gave a score above 0 for 58\% of the non ideal scenes with markers, contrary to an average of 11\% for the other subjects.

@@

In this section, we study the scores given by the subject while listening to the several types of scenes, namely the ideal scenes with markers, non ideal scenes with markers, ideal scenes without markers and non ideal scenes without markers (Figure~\ref{fig:xp2_Ascene}). The repeated measure ANOVA applied to $\mathcal{A}_{subject}$ shows a significant effect of the type environment (Type: $F[1,10]=175$, $p<0.01$), of the presence/absence of markers (marker: $F[1,10]=7$, $p<0.05$), and of the interaction between those two factors ($F[1,10]=67$, $p<0.01$).

@@

The \emph{post hoc} analysis exhibits significant differences between all groups of observations, notably between the ideal scenes with markers and the ideal scenes without markers ($p<0.05$) as well as between the non ideal scenes with markers and the non ideal scenes without markers ($p<0.01$).

@@

The significant interaction shows that the type of environment influences the effect of the presence/absence of the markers. Indeed, the average difference between the scenes with markers and the scenes without markers is larger for the non ideal scenes ($1.1$) than for the ideal scenes ($0.5$).
